---
title: "An Overview of COVID 19 Progression in Four Major Countries"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{An Overview of COVID 19 Progression in Four Major Countries}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE,
  warning = FALSE,
  fig.width = 7,
  fig.height = 5
)
```

```{r setup, include=FALSE}
library(covtrack)
library(shiny)
library(shinydashboard)
library(dplyr)
library(ggplot2)
library(plotly)
library(leaflet)
library(sf)
library(lubridate)
library(tidyr)
library(stringr)

Aus_cases <- system.file("ext", "Aus_cases.csv", package="covtrack")
Aus_deaths <- system.file("ext", "Aus_deaths.csv", package="covtrack")
Aus_deaths <- system.file("ext", "Aus_state_codes.xlsx", package="covtrack")
Aus_deaths <- system.file("ext", "Aus_deaths.csv", package="covtrack")
Aus_state_codes <- system.file("ext", "Aus_state_codes.xlsx", package="covtrack")
Brazil_state_codes <- system.file("ext", "Brazil_state_codes.xlsx", package="covtrack")
Brazil_states <- system.file("ext", "Brazil_states.xlsx", package="covtrack")
covid_stringency_index <- system.file("ext", "covid-stringency-index.csv", package="covtrack")
Death_rate <- system.file("ext", "Death_rate_null.png", package="covtrack")
India_latlong <- system.file("ext", "India_latlong.xlsx", package="covtrack")
India_state_codes <- system.file("ext", "India_state_codes.csv", package="covtrack")
India_states <- system.file("ext", "India_states.csv", package="covtrack")
New_cases <- system.file("ext", "New_cases.png", package="covtrack")
New_deaths <- system.file("ext", "New_deaths.png", package="covtrack")
us_state_codes <- system.file("ext", "us_state_codes.csv", package="covtrack")
US_states <- system.file("ext", "US_states.csv", package="covtrack")
```


```{r loader, include=FALSE}
#india <- read.csv(here::here("data", "India_states.csv"))
india <- read.csv(India_states)

india1 <- india %>% pivot_longer(cols = 3:41, names_to = "state", values_to = "value")

india2 <- india1 %>% pivot_wider(names_from = Status, values_from = value)

india2 <- india2 %>% rename(State_code = state)

# isc <- read.csv(here::here("data", "India_state_codes.csv"))
isc <- read.csv(India_state_codes)

india3 <- india2 %>% left_join(isc, by = "State_code")

# US <- read.csv(here::here("data", "US_states.csv"))
US <- read.csv(US_states)

# brazil <- readxl::read_xlsx(here::here("data", "Brazil_states.xlsx"))
brazil <- readxl::read_xlsx(Brazil_states)

# aus_cases <- read.csv(here::here("data", "Aus_cases.csv"))
aus_cases <- read.csv(Aus_cases)

aus_cases1 <- aus_cases %>% pivot_longer(cols = 2:9, names_to = "state", values_to = "new_cases")
aus_cases2 <- aus_cases1 %>% rename(date = `ï..Date`)

# aus_deaths <- read.csv(here::here("data", "Aus_deaths.csv"))
aus_deaths <- read.csv(Aus_deaths)

aus_deaths1 <- aus_deaths %>% pivot_longer(cols = 2:9, names_to = "state", values_to = "new_deaths")
aus_deaths2 <- aus_deaths1 %>% rename(date = `ï..Date`)

aus_join <- aus_cases2 %>% left_join(aus_deaths2, by = c("date" = "date", "state" = "state"))

aus_join[is.na(aus_join)] = 0

aus_join$date <- paste0(aus_join$date, "-2020")

aus_join$date <- dmy(aus_join$date)
aus_join <- aus_join %>% rename(State_code = state)

# aus_ll <- readxl::read_xlsx(here::here("data", "Aus_state_codes.xlsx"))
aus_ll <- readxl::read_xlsx(Aus_state_codes)

aus_ll <- aus_ll %>% rename(state = State)

aus_final <- aus_join %>% left_join(aus_ll, by = "State_code")

india3$Date <- dmy(india3$Date)

# india_ll <- readxl::read_xlsx(here::here("data", "India_latlong.xlsx"))
india_ll <- readxl::read_xlsx(India_latlong)


india_final <- india3 %>% left_join(india_ll, by = "State")

india_final <- india3 %>% filter(State != "State Unassigned") %>% filter(State != "Total")

# us_ll <- read.csv(here::here("data", "us_state_codes.csv"))
us_ll <- read.csv(us_state_codes)


US <- US %>% rename(State_code = state)

us_final <- US %>% left_join(us_ll, by = "State_code")

us_final$submission_date <- mdy(us_final$submission_date)

us_final <- us_final[!is.na(us_final$Latitude), ]

brazil1 <- brazil %>% filter(state != "TOTAL")

# brazil_ll <- readxl::read_xlsx(here::here("data", "Brazil_state_codes.xlsx"))
brazil_ll <- readxl::read_xlsx(Brazil_state_codes)

brazil1 <- brazil1 %>% rename(State_codes = state)

brazil2 <- brazil1 %>% left_join(brazil_ll, by = "State_codes")

brazil_final <- brazil2 %>% select(date, country, State_codes, newDeaths, newCases, totalCases, state)

brazil_final <- brazil_final %>% rename(State_code = State_codes)
brazil_final1 <- brazil_final %>% select(date, State_code, state, country, newCases, newDeaths)

aus_final$country <- "Australia"

aus_final <- aus_final %>% rename(newCases = new_cases, newDeaths = new_deaths)
aus_final1 <- aus_final %>% select(date, State_code, state, country, newCases, newDeaths)

us_final$country <- "USA"

us_final <- us_final %>% rename(date = submission_date, state = State, newCases = new_case, newDeaths = new_death)
us_final1 <- us_final %>% select(date, State_code, state, country, newCases, newDeaths)

india_final$country <- "India"

india_final <- india_final %>% rename(date = Date, newCases = Confirmed, newDeaths = Deceased, state = State)
india_final1 <- india_final %>% select(date, State_code, state, country, newCases, newDeaths)

final <- rbind(us_final1, india_final1, brazil_final1, aus_final1)

final <- final

brazil_states <- geojsonio::geojson_read("https://raw.githubusercontent.com/codeforamerica/click_that_hood/master/public/data/brazil-states.geojson", what = "sp")


brazil_states1 <- st_as_sf(brazil_states)

brazil_states1 <- brazil_states1 %>% rename(state = name)

brazil_states1 <- brazil_states1 %>% rename(State_code = sigla)

brazil_states2 <- brazil_states1 %>% select(State_code, geometry)

brazil_join <- brazil_final1 %>% left_join(brazil_states2, by = "State_code")

us_states <- geojsonio::geojson_read("https://raw.githubusercontent.com/PublicaMundi/MappingAPI/master/data/geojson/us-states.json", what = "sp")

us_states1 <- st_as_sf(us_states)

us_states1 <- us_states1 %>% rename(state = name)

us_states2 <- us_states1 %>% select(state, geometry)

us_join <- us_final1 %>% left_join(us_states2, by = "state")

india_states <- geojsonio::geojson_read("https://raw.githubusercontent.com/geohacker/india/master/state/india_telengana.geojson", what = "sp")

india_states1 <- st_as_sf(india_states)

india_sort <- india_final1 %>% select(State_code, state)

india_sort1 <- india_sort %>% mutate(ID = row_number())

india_sort1 <- india_sort1 %>% filter(ID <= 36)

india_states2 <- india_states1 %>% rename(ID = ID_1)

india_sort3 <- india_sort1 %>% left_join(india_final1, by = "state")

india_states2 <- india_states2 %>% select(ID, geometry)

india_final2 <- india_sort3 %>% left_join(india_states2, by = "ID")

india_final2 <- india_final2 %>% select(date, State_code.y, state, country, newCases, newDeaths, geometry)
india_final2 <- india_final2 %>% rename(State_code = State_code.y)

australia_states <- geojsonio::geojson_read("https://raw.githubusercontent.com/rowanhogan/australian-states/master/states.geojson", what = "sp")

australia_states1 <- st_as_sf(australia_states)

australia_states1 <- australia_states1 %>% rename(state = STATE_NAME)

australia_states1 <- australia_states1 %>% select(state, geometry)

australia_join <- aus_final1 %>% left_join(australia_states1, by = "state")


final <- rbind(us_join, india_final2, brazil_join, australia_join)
final <- st_as_sf(final)

final <- final %>% filter(date <= "2020-09-25")

final$newCases <- str_remove_all(final$newCases, "-")
final$newCases <- as.double(final$newCases)

final$newDeaths <- str_remove_all(final$newDeaths, "-")
final$newDeaths <- as.double(final$newDeaths)


cd <- final

cd <- cd %>% mutate(month = substr(date, 6, 7))

cd <- cd %>% group_by(state, month) %>% mutate(m_case = sum(newCases)) %>% mutate(m_death = sum(newDeaths))

cd <- cd %>% mutate(cbd = (m_death/m_case)*100)

cd$cbd[is.nan(cd$cbd)]<-0

cd$cbd <- round(cd$cbd, 2)

cd <- cd %>% group_by(country, month) %>% mutate(countr_mcase = sum(newCases)) %>% mutate(countr_mdeath = sum(newDeaths))

cd <- cd %>% mutate(countr_cbd = (countr_mdeath/countr_mcase)*100)

cd$countr_cbd <- round(cd$countr_cbd, 2)

# stringent <- read.csv(here::here("data", "covid-stringency-index.csv"))
stringent <- read.csv(covid_stringency_index)

stringent <- stringent %>% filter(Date >= "2020-01-22" & Date <= "2020-09-25")

stringent <- stringent %>% filter(Code %in% c("USA", "AUS", "BRA", "IND"))

stringent <- stringent %>% mutate(month = substr(Date, 6, 7)) %>%
    group_by(month, Entity) %>%
    mutate(m_string = mean(Stringency.Index..OxBSG.))

stringent$m_string <- round(stringent$m_string, 2)

stringent <- stringent %>% rename(country = Entity)

stringent <- stringent %>% select(country, month, m_string)

stringent$country[stringent$country == "United States"] <- "USA"

stringent <- distinct(stringent)
cd <- cd %>% left_join(stringent)

cd$month <- as.numeric(cd$month)

cd1 <- cd %>% select(month, country, state, m_case, m_death, m_string)
cd1 <- as.data.frame(cd1)
cd1 <- distinct(cd1)

cd2 <- cd %>% select(month, country, state, cbd)
cd2 <- as.data.frame(cd2)
```

This package provides the shiny app to visualise the progression of COVID 19 with time in four major countries affected - `r unique(final$country)`. Visualisation is done for 4 major aspects - new cases, deaths, death rate with respect to cases and stringency index (level of restrictions imposed by governments to stop the virus).


Examples of visualisations included in covtrack:

1. New cases: An example of new cases visualised after the user selects USA as the country and 22-06-2020 as the date on the time slider. Both options provided in the app to choose from:
```{r mapper}
final_filt <- final %>% 
  filter(country == "USA") %>% filter(date == "2020-06-22")
labels <- sprintf("<strong>%s</strong><br/>%g new cases",
                          final_filt$state, final_filt$newCases) %>% lapply(htmltools::HTML)

bins <- c(0, 50, 100, 200, 400, 600, 800, 1000, 1200, 1400, 1600, 1800, 2000, Inf)
beaut <- colorBin("YlOrRd", domain = final_filt$newCases, bins = bins)

map_USA <- leaflet(final_filt) %>%
  setView(-110, 52.8, 2.5) %>%
  addTiles("MapBox") %>%
  addPolygons(fillColor = ~beaut(final_filt$newCases),
              weight = 2,
              opacity = 1,
              color = "white",
              dashArray = "3",
              fillOpacity = 1,
              highlight = highlightOptions(weight = 5,
                                           color = "#600",
                                           dashArray = "",
                                           fillOpacity = 0.7,
                                           bringToFront = TRUE),
              label = labels)%>% addLegend(pal = beaut,
                                           values = ~newCases, 
                                           opacity = 0.7, title = NULL, 
                                           position = "topright")

map_USA
```


1. Death rate: This is calculated as the percentage of monthly deaths out of monthly cases in a particular country or state in one of those countries. Here is what that looks like for all four countries included in the analyses for this app:

```{r dr}
dr <- cd %>%
  ggplot(aes(x = as.factor(month), y = countr_cbd, text = paste0("Country: ", country, ",Death rate: ", cbd, "  Month: ", month))) +
            geom_line(aes(color = country, group = country)) +
            labs(x = "Months",
                 y = "Death rate per monthly cases") +
            theme_bw() +
            theme(legend.title = element_blank())

ggplotly(dr)
```

